---
title: "DSA2101 Midterms"
author: "Michelle Yong Pei Feng A0221827H"
date: "10/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(ggplot2)
library(lubridate)
library(sf)
```

```{r data, include=TRUE}
#Question 2.1
resale1 <- read.csv("../data/resale1.csv") %>% as_tibble()
resale2 <- read.csv("../data/resale2.csv") %>% as_tibble()

resale2 <- resale2 %>% 
  mutate(remaining_lease = as.integer(str_extract(remaining_lease,"^.{2}"))) %>%
  mutate(month = ifelse(month == "2020-Aug", "2020-08", month))
  
resale_prices <- full_join(resale1, resale2) %>%
  mutate(town = toupper(town))

#Question 2.2
resale_prices <- resale_prices %>%
  mutate(price_psm = resale_price/floor_area_sqm,
         lstorey = as.integer(str_extract(storey_range,"^.{2}")),
         ustorey = as.integer(str_extract(storey_range, ".{2}$"))) %>%
  separate(month, c("year", "month"), sep="-") %>%
  mutate(month = as.numeric(month), 
         year = as.numeric(year),
         town = ifelse(town=="KALLANG/WHAMPOA", "CENTRAL AREA", town))

#Question 2.3
CPI_csv <- read.csv("../data/CPI_data.csv", skip=10, head=TRUE) %>% 
  as_tibble()

CPI_data <- slice_head(CPI_csv, n=1) %>%
  rownames_to_column() %>% 
  select(-Data.Series) %>%
  mutate(rowname = "CPI") %>%
  gather("month", "CPI", -rowname) %>% 
  spread(rowname, "CPI") %>%
  mutate(year = as.integer(str_extract(month,"[[:digit:]]{4}"))) %>%
  mutate(month = match(str_sub(month,-3,-1), month.abb),
         CPI = as.numeric(CPI)) %>%
  filter(year>2014) %>%
  arrange(year, month)

resale_prices <- left_join(resale_prices, CPI_data, by = c("year", "month")) %>%
  mutate(price_psm_adj = CPI/100*price_psm)

median_prices <- resale_prices %>%
  filter(!is.na(price_psm_adj), flat_type == "4 ROOM") %>%
  group_by(year, month) %>%
  summarize(median_adj = median(price_psm_adj),
            median_unadj = median(price_psm)) %>%
  mutate(date = make_date(year,month))

ggplot(median_prices, aes(x=date)) + 
  geom_line(aes(y=median_adj, color="black")) +
  geom_line(aes(y=median_unadj, color="grey"), color="grey") +
  scale_color_manual(name="Legend", values = c("Adjusted"="black", 
                                               "Unadjusted"="grey")) +
  theme(legend.position = c(0.1,0.80)) +
  labs(title="4-Room HDB Resale Prices") +
  xlab("Month") +
  ylab("Price")
```

Generally, the adjusted and unadjusted prices follow the same upward trend
closely. The adjusted median prices for 4-room flats before
the base year 2019 are lower than the unadjusted prices. The prices were
similar throughout 2019 and beginning of 2020. Adjusted prices began to rise
faster than the unadjusted prices in May 2020 approximately.The growing
distance between the 2 lines suggests that inflation rates are rising.

```{r, include=TRUE}
#Question 2.4
med_price_town_tbl <- resale_prices %>%
  filter(year==2020 | (year==2021 & month<9), flat_type == "4 ROOM") %>%
  group_by(town) %>%
  summarize(median_adj = median(price_psm_adj))

#Question 2.5
plm <- function(x,y){
  combn_x <- combn(x,2)
  combn_y <- combn(y,2)
  diff_x <- apply(combn_x,2,function(x){x[2]-x[1]})
  diff_y <- apply(combn_y,2,function(x){x[2]-x[1]})
  gradient <- diff_y/diff_x
  gradient[!is.finite(gradient)] <- NA
  b_hat <- median(gradient, na.rm = TRUE)
  a_hat <- median(combn_y-b_hat*combn_x, na.rm = TRUE)
  return(c(b_hat, a_hat))
}

#Question 2.6
area_prices <- resale_prices %>%
  filter(year==2020 | (year==2021 & month<9)) %>%
  mutate(date = make_date(year,month)) %>%
  group_by(town, date) %>%
  mutate(med_price_psm_adj = median(price_psm_adj)) %>%
  select(town, med_price_psm_adj, date) %>%
  distinct() 

plm_tbl <- area_prices %>%
  group_by(town) %>%
  summarize(plm_b = plm(date, med_price_psm_adj)[1],
            plm_a = plm(date, med_price_psm_adj)[2])

greatest_rate <- slice_max(plm_tbl, plm_b, n=3)
lowest_rate <- slice_min(plm_tbl, plm_b, n=3)
```

Queenstown, Choa Chu Kang and Central area had the greatest rate of increase
in monthly median adjusted price, while Bukit Timah, Bukit Merah, and SengKang
had the lowest rates.

```{r, include=TRUE}
#Question 2.7
model_prices <- resale_prices %>%
  filter(year>2020) %>%
  group_by(flat_model, town) %>%
  count() %>%
  pivot_wider(id_cols=town, names_from=flat_model, values_from=n) %>% 
  ungroup()
```

Examples of exclusive flat models include the Improved-Maisonette exclusive 
to Hougang, Type S1 and S2 flats exclusive to central areas. There are other
models such as the Premium Apartment Loft, Premium Maisonette and Multi
Generation flats limited to 2 or lesser towns.

```{r, include=TRUE}
#Question 2.8
pln_areas <- readRDS("../data/sg_planning_areas.rds")
central <- c("KALLANG", "NOVENA", "TANGLIN", "DOWNTOWN CORE", "MUSEUM", 
             "NEWTON", "ORCHARD", "OUTRAM", "RIVER VALLEY", "ROCHOR",
             "SINGAPORE RIVER")
pln_areas <- filter(pln_areas, pln_areas$PLN_AREA_N %in% central) %>% 
  st_union() %>%
  st_union(pln_areas,.)
ggplot(pln_areas) +
  geom_sf()

#Question 2.9
# town_prices <- resale_prices %>%
#   filter(year==2015, flat_type=="4 ROOM") %>%
#   group_by(town) %>%
#   summarize(med_price_psm_adj=median(price_psm_adj)) %>%
#   left_join(pln_areas, by=c("town"="PLN_AREA_N"))
# 
# st_touches(town_prices$geometry, st_geometry(a))
```

Question 2.10\
We may associate the growing demand for certain housing with the
percentage increase in resale prices after every year. By analyzing
the percentage change in prices for all areas in a year, we can plot a graph
of percentage increase against floor area to linearize the relationship. If
the plot shows a upward trend, we may conclude that there is greater demand
for bigger houses. Similarly, plotting the graphs of percentage changes in
price for central vs non-central areas in a given year can show which housing
locations are more popular.