---
title: "DSA2101 Tutorial 7"
runningheader: "Raking and clustering implementation" # only for pdf output
subtitle: "Raking and clustering implementation" # only for html output
author: "Michelle Yong Pei Feng (A0221827H)"
date: "15 November 2021"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
library(readxl)
library(tidyverse)
library(survey)
library(cluster)
library(gridExtra)
library(factoextra)
```

## Survey Background 

The Questionnaire
1. I don’t feel particularly pleased with the way I am. (R)   
2. I am intensely interested in other people.   
3. I feel that life is very rewarding.   
4. I have very warm feelings towards almost everyone.   
5. I rarely wake up feeling rested. (R)   
6. I am not particularly optimistic about the future. (R)   
7. I find most things amusing.   
8. I am always committed and involved.   
9. Life is good.   
10. I do not think that the world is a good place. (R) _  
11. I laugh a lot.   
12. I am well satisfied about everything in my life.   
13. I don’t think I look attractive. (R)   
14. There is a gap between what I would like to do and what I have done. (R)   
15. I am very happy.   
16. I find beauty in some things.   
17. I always have a cheerful effect on others.   
18. I can fit in (find time for) everything I want to.   
19. I feel that I am not especially in control of my life. (R)   
20. I feel able to take anything on.   
21. I feel fully mentally alert.   
22. I often experience joy and elation.   
23. I don’t find it easy to make decisions. (R)   
24. I don’t have a particular sense of meaning and purpose in my life. (R)   
25. I feel I have a great deal of energy.   
26. I usually have a good influence on events.   
27. I don’t have fun with other people. (R)   
28. I don’t feel particularly healthy. (R)   
29. I don’t have particularly happy memories of the past. (R)

Happiness score = Sum of scores divided by 29.

## Data Manipulation

For formatting, column names of question responses have been converted to abbreviations,
such as "I don’t feel particularly pleased with the way I am." to "Q1" and 
"I am intensely interested in other people." to "Q2" etc. for all 29 survey
questions. Items marked (R) with scores reversed are also corrected for the happiness
count.

```{r data1, cache=TRUE, warning=FALSE}
#Data manipulation
happy <- read_excel("../data/happ_svy_data.xlsx") %>%
  as_tibble()
names <- paste0("Q",1:29)
colnames(happy) <- c(names, "citizenship", "residential","gender")

#Adding happiness count for each row
reverse <- c(1,5,6,10,13,14,19,23,24,27,28,29)
order <- c(1:29) %>%
  setdiff(reverse)
happy <- happy %>%
  mutate_at(order, funs(recode(.,`Strongly disagree`=1, `Moderately disagree`=2,   
                                `Slightly disagree`=3, `Slightly agree`=4, 
                                `Moderately agree`=5, `Strongly agree`=6))) %>%
  mutate_at(reverse, funs(recode(.,`Strongly disagree`=6, `Moderately disagree`=5,   
                               `Slightly disagree`=4, `Slightly agree`=3, 
                               `Moderately agree`=2, `Strongly agree`=1))) %>%
  mutate(happiness=rowSums(across(Q1:Q29))/29)

happy
```

## Removing Outlier

```{r fig1, fig.margin=TRUE, fig.cap = "Happiness boxplot with outlier", cache=TRUE}
ggplot(happy) +
  geom_boxplot(aes(citizenship, happiness)) +
  geom_boxplot(aes(gender, happiness)) +
  geom_boxplot(aes(residential, happiness)) +
  scale_x_discrete(limits=c("M","F","Foreigner","Singaporean/PR","On campus","Off campus")) +
  geom_vline(xintercept = c(2.5,4.5), linetype="dashed", size=0.6) +
  labs(x="Demogrpahic", y="Happiness Count", 
       title="Happiness boxplot for each demographic group before raking")
```

The outlier as seen from the boxplot belongs to the demographic: Female, foreigner and off-campus.
Since the observed value deviates from the majority drastically, the outlier is removed 
for better data analysis.

```{r fig2, fig.cap = "Happiness boxplot without outlier", cache=TRUE}
#Removing outlier
outlier <- happy %>% 
  filter(happiness<2)
outlier
happy <- happy %>%
  filter(happiness>2)

ggplot(happy) +
  geom_boxplot(aes(citizenship, happiness)) +
  geom_boxplot(aes(gender, happiness)) +
  geom_boxplot(aes(residential, happiness)) +
  scale_x_discrete(limits=c("M","F","Foreigner","Singaporean/PR","On campus","Off campus")) +
  geom_vline(xintercept = c(2.5,4.5), linetype="dashed", size=0.6) +
  labs(x="Demogrpahic", y="Happiness Count", 
       title="Happiness boxplot for each demographic group before raking")
```

The final box plot shows that generally, females, foreigners, and students on campus 
have higher happiness scores than their counterparts. A possible prediction may be that 
most foreign students are staying on campus and consists of more females than males.

## Raking: Correction for Non-response

There are 44 males and 28 females in the sample population, while there are
65 males and 40 females in the entire population. Hence, weighing the 
responses with raking by gender corrects for non-response in the survey data.

```{r data2(i), cache=TRUE, warning=FALSE}
#Creating survey object
happy$ID <- seq.int(nrow(happy))
happy$fpc <- 105
happy_data <- happy %>%
  svydesign(id=~ID, data=., fpc=~fpc)

#Before raking
svymean(~happiness, design=happy_data)
confint(svymean(~happiness, design=happy_data))
```

```{r data2(ii), cache=TRUE, warning=FALSE}
#Raking by gender
happy_pop <- data.frame(gender=c("M","F"),Freq=c(65,40))

happy_data.raked <- rake(happy_data, sample=list(~gender),
                         population=list(happy_pop))
#After raking
svymean(~happiness, design=happy_data.raked)
confint(svymean(~happiness, design=happy_data.raked))
```

After raking, the happiness mean decreased from 3.6825 to 3.6806, with a smaller 
standard deviation from 0.0392 to 0.0384. The confidence interval of the raked 
data also narrowed. 


## Ward's Clustering

To analyse the correlation between each individual question scores, Ward's 
hierarchical clustering is utilized to group questions will higher correlation into
the same cluster.

```{r data3, cache=TRUE, warning=FALSE}
happy_questions <- happy %>%
  group_by(Q1)

cor_happy <- happy %>%
  select(Q1:Q29) %>%
  select(.,which(!sapply(., is.factor))) %>%
  cor(., use="pair")

cor_happy_df <- as.data.frame(cor_happy,row.names=NULL) %>%
  mutate(var1=row.names(cor_happy)) %>%
  pivot_longer(Q1:Q29, names_to ="var2",values_to ="correlation")

hc <- hclust(as.dist((1-cor_happy)/2), method="ward.D")
ord <- order.dendrogram(as.dendrogram(hc))
cor_happy_df2 <- mutate(cor_happy_df,
                          var1=factor(var1,levels=row.names(cor_happy)[ord]),
                          var2=factor(var2,levels=row.names(cor_happy)[ord]))
```

```{r fig3, fig.cap = "Happiness Dendrogram with Ward's Clustering", cache=TRUE}
plot(as.dendrogram(hc))

ggplot(cor_happy_df2) +
  geom_tile(aes(x=var1,y=var2,fill=correlation)) +
  scale_fill_gradient(low="white", high="darkblue") +
  theme(axis.text = element_text(size=6)) +
  labs(x="Response 1", y="Response 2", title="Correlation Heat Map for the 29 Survey Questions")
```

From the correlation heatmap, there is strong association between questions within the cluster
containing Q12, Q22, Q9 and Q15. These questions belongs to the questions: 

Q12:  "I am well satisfied about everything in my life."
Q22:  "I often experience joy and elation."
Q9:  "Life is good."
Q15:  "I am very happy."

The questions derived similar results given a person's happiness level, hence they 
have similar scores and higher correlation as compared to other questions. For example, 
a happy person is more likely score well for all 4 questions, but an unhappy question 
will score poorly. 

Alternatively, Q2 and Q25 have lower correlation with all other questions. These 
questions are:

Q2:  "I am intensely interested in other people."
Q25:  "I feel I have a great deal of energy."

These two questions have poor correlation will all other questions, hence their scores 
are not associated to the other questions. A possible reason maybe that Q2 and Q25 have
poor relation to one's happiness, resulting in varying responses.

## Kmeans Clustering

Kmeans clustering can be used to study questions which are more likely to score well
or poorly together, as well as the effect of the cluster on the overall happiness count.

```{r fig4, fig.cap = "Happiness Kmeans Clustering with Different K Values", cache=TRUE}
happy1 <- happy %>%
  select(Q1:Q29, happiness, ID)

k2 <- kmeans(happy1, centers = 2, nstart = 25)
k3 <- kmeans(happy1, centers = 3, nstart = 25)
k4 <- kmeans(happy1, centers = 4, nstart = 25)
k5 <- kmeans(happy1, centers = 5, nstart = 25)

p1 <- fviz_cluster(k2, geom = "point", data = happy1) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = happy1) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = happy1) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = happy1) + ggtitle("k = 5")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

The representation shows the distribution of points for each k value. To find the 
optimal k value, we can find the k value with the minimum cluster gap.

```{r fig5, fig.margin=TRUE, fig.cap = "Optimal K Value", cache=TRUE}
set.seed(123)
gap_stat <- clusGap(happy1, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```

From the margin figure, the optimal k value is 8 as the cluster gap is minimal. By
using the optimal 8 value, we can group the survey questions and study their impact 
on a student's happiness count.

```{r fig6, fig.cap = "Kmeans Clustering at Optimal K value", cache=TRUE}
final <- kmeans(happy1, 8, nstart = 25)
fviz_cluster(final, geom = "point",  data = happy1) + ggtitle("k = 8")
```

```{r data4, cache=TRUE, warning=FALSE}
happy_kmeans <- happy1 %>%
  mutate(cluster=final$cluster) %>%
  group_by(cluster) %>%
  summarise_all("mean") %>%
  pivot_longer(cols=c(Q1:Q29, happiness),
               names_to="value", values_to="mean") %>%
  mutate(h=ifelse(value=="happiness", 1, 0))
```

```{r fig7, fig.cap = "Kmeans Clustering, k=8", cache=TRUE}
ggplot(happy_kmeans, aes(reorder(value,-mean), mean, group=cluster)) +
  geom_point(aes(color=h)) +
  facet_wrap(~ cluster) +
  theme_bw() +
  
  theme(axis.text.x=element_text(angle=70, hjust=1, size=5),
        legend.position="none") +
  labs(x="Survey Catagories", y="Mean Value", 
       title="Mean Values For Each Survey Catagory Within Each Cluster")
```

Each graph belongs to each cluster of survey questions and their mean values are tabulated. 
The graphs show the general relationship between the questions and the overall 
happiness score. 

The representation shows the variation of mean scores and the proportion of points 
above and below the happiness count for each cluster. For example, cluster 2 has 
all points above score 3, excluding Q25 having 2. Therefore, we can predict that for a student 
who scores well for most questions, he/she is more likely to get a 
score of 2 for Q25 and a happiness count of approximately 4.7. 
The clusters help group students will similar responses and estimate 
their happiness score. 



