---
title: "DSA2101 Tutorial 6"
author: "Michelle Yong Pei Feng A0221827H"
date: "10/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Hmisc)
library(readxl)
library(wordcloud)
library(tm)
```

```{r data, include = TRUE}
artist_info <- read.csv("../data/artist_data.csv") %>% as_tibble()
artist_work <- read.csv("../data/artwork_data.csv", skip=1, header=FALSE) %>% 
  as_tibble()
col_names <- readLines("../data/artwork_data.csv", n=1)
col_names <- str_split(col_names, ",")[[1]]
colnames(artist_work) <- col_names

#Convert to numeric
artist_work$year <- as.numeric(artist_work$year)

# Warning message:
# NAs introduced by coercion
# Some year data was missing, missing values replaced by NA.

#Place of Birth
artists_nation <- read_xlsx("../data/artists_nation.xlsx")

x <- artist_info %>% 
  filter(placeOfBirth != "") %>%
  separate(placeOfBirth, c("city", "birth_area"), ", ", extra="merge") %>%
  left_join(artists_nation, by="birth_area") %>%
  select(ï..id, nation)

artist_info <- artist_info %>% left_join(x, by="ï..id")

#No information
no_info <- anti_join(artist_work, artist_info, by=c("artist"="name")) %>%
  select(artist, artistId) %>%
  unique()
no_info

#Remove an Artist Outlier
most_work <- artist_work %>% group_by(artist) %>%
  count() %>%
  arrange(desc(n)) %>% ungroup() %>%
  slice(1)
most_work
artist_work <- anti_join(artist_work, most_work, by="artist")

#Gender counts By Century
work2 <- mutate(artist_work, century=paste0(acquisitionYear %/% 100, "th C."))
genderByCentury <- left_join(work2, artist_info, by=c("artist"="name")) %>%
  filter(century!="NAth C.") %>%
  filter(gender=="Male" | gender=="Female") %>%
  group_by(century, gender) %>% 
  count() %>%
  pivot_wider(id_cols=gender, names_from=century, values_from=n)
genderByCentury

#Popularity of Medium
#1
tokens <- str_split(artist_work$medium, boundary("word")) %>%
  unlist() %>% str_to_lower()
#2
tokens_no_stop <- removeWords(tokens,
                              c(stopwords(), "paper", "canvas")) %>%
  Filter(function(x) nchar(x) >0, .)
#3
tmp_table <- table(tokens_no_stop)
ww <- names(tmp_table)
vv <- as.numeric(tmp_table)
set.seed(2101)
wordcloud(ww, vv, scale=c(1.5,.3), min.freq = 30)
```

1. Sentences in each row of the medium column are broken down into individual
   words by str_split() and converted to lower caps by str_to_lower().
2. removeWords() from the 'tm' library converts filler words like "on", "and" 
   to "". "paper" and "canvas" were also converted as they appear too 
   frequently and are not meaningful. "" are removed with the filter function.
3. Names and frequencies of the different mediums are assigned to ww and vv. 
   wordcloud() presents that the top mediums used are "paint", "graphite"
   and "watercolour".


```{r, include = TRUE}
#Versatile artists
versatile_artists <- artist_work %>%
  group_by(artist) %>%
  filter(n()>9) %>%
  filter(n_distinct(medium)==n()) %>%
  ungroup() %>%
  select(artist, medium, year)
versatile_artists

#Popular artists
popular_artists <- artist_work %>% 
  mutate(difference = acquisitionYear-year) %>%
  group_by(artistId) %>%
  filter(n()>9, min(year)>1969) %>%
  summarise(median_yrs = median(difference), .groups="drop") %>%
  filter(median_yrs<=1) %>%
  select(artistId, median_yrs)
popular_artists
```

To Explore\
From the github link, art works acquired by Tate were grouped according to 
world continents. The data shows that most of the work were from Europe.
(Tate Collection Geolocalized by Corentin Cournac, Mathieu Dauré, William 
Duclot and Pierre Présent.) 
It would be interesting to group the acquired art works by countries within
Europe itself and study the data against the placeOfBirth/placeOfDeath of the 
artists.


